%#!lualatex
\documentclass[lualatex,a4paper,11pt]{jlreq}

%%%%%%%%%%%%%%%%
% additional packages
\usepackage{amsmath}
\usepackage{array}
\usepackage[defaultsups]{newpxtext}
\usepackage[zerostyle=c,straightquotes]{newtxtt}
\usepackage{newpxmath}
\usepackage[svgnames]{xcolor}
\usepackage[hyperfootnotes=false]{hyperref}
\usepackage{hologo}
\usepackage{xspace}

\def\Pkg#1{\textsf{#1}}
\def\Opt#1{\texttt{#1}}
\def\ClsY#1{\Pkg{\textcolor{Black}{#1}}}
\def\ClsT#1{\Pkg{\textcolor{Green}{#1}}}
\def\ClsYT#1{\Pkg{\textcolor{Blue}{#1}}}
\def\file#1{\textsf{#1}}
\def\code#1{\texttt{#1}}

\usepackage[english]{babel}% References in English

\usepackage{shortvrb}
\MakeShortVerb*{|}
%%%%%%%%%%%%%%%%

% logos
\def\eTeX{$\varepsilon$-\TeX}
\def\pTeX{p\kern-.10em\TeX}
\def\upTeX{u\pTeX}
\def\epTeX{$\varepsilon$-\pTeX}
\def\eupTeX{$\varepsilon$-\upTeX}
\def\pLaTeX{p\LaTeX}\def\pLaTeXe{p\LaTeXe}
\def\upLaTeX{u\pLaTeX}\def\upLaTeXe{u\pLaTeXe}
\def\pdfTeX{\hologo{pdfTeX}}
\def\pdfLaTeX{\hologo{pdfLaTeX}}
\def\XeTeX{\hologo{XeTeX}}
\def\XeLaTeX{\hologo{XeLaTeX}}
\def\LuaTeX{\hologo{LuaTeX}}
\def\LuaLaTeX{\hologo{LuaLaTeX}}

\def\TL{\TeX\ Live\xspace}
\def\JTDC{Japanese \TeX\ Development Community}

\def\_{\leavevmode\vrule width .45em height -.2ex depth .3ex\relax}

\frenchspacing
\begin{document}

\title{\textsf{\textbf{Guide to Japanese typesetting with \LaTeX}}}
\author{\JTDC\null
\thanks{\url{https://texjp.org},\ e-mail: \texttt{issue(at)texjp.org}}}
\maketitle

% 日本語組版は西欧言語とは全然違う
Typesetting requirements of Japanese are completely different from
those of Western languages, so setting parameter values and additional
hyphenation patterns are not enough.

% babel-japanese (japanese.ldf) は訳語を登録するだけ
% 日本語組版に必要な設定は行わない
The package \Pkg{babel-japanese} provides ``\Opt{japanese}''
option for Babel package.
However, it provides only translated replacement texts for keywords
and dates (e.g. ``Table of contents'' $\rightarrow$ ``目次'',
\verb+\today+ $\rightarrow$ ``\和暦\today'')
and it does not care about any Japanese typesetting requirements.


% とりあえず LuaLaTeX + \usepackage{luatexja} オススメ
\section{Short introduction: using \LuaLaTeX}

To obtain a properly formatted Japanese document, the easiest way
is to use \LuaLaTeX\ with Lua\TeX-ja package.

\begin{itemize}
  \item Using \Pkg{jlreq} class:
\begin{verbatim}
    \documentclass[book]{jlreq}
    \begin{document}
    \chapter{日本語の文書}% Japanese document
    こんにちは，日本。% Hello, Japan.
    \end{document}
\end{verbatim}
    Save the above as \file{jp1.tex} and run a command.
\begin{verbatim}
    $ lualatex jp1
\end{verbatim}
    You will get \file{jp1.pdf} which is properly formatted and
    a Japanese font family ``HaranoAji Mincho/Gothic'' is embedded.
 \item If you prefer Western classes, you can load \Pkg{luatexja} directly:
\begin{verbatim}
    \documentclass{book}
    \usepackage{luatexja}
    \begin{document}
    ...
\end{verbatim}
\end{itemize}


% 他の選択肢：upLaTeX が先
\section{Relatively stable alternative: \upLaTeX}

\LuaLaTeX\ is relatively new,
and Lua\TeX-ja is under active development.
As a more stable alternative, you can use \upLaTeX.
It is a somewhat legacy implementation, which requires TFM files
for typesetting Japanese characters (called JFM).
If you use only two fixed-width Japanese fonts, using a standard
class is enough:
\begin{verbatim}
    \documentclass[uplatex]{jsbook}% or, \documentclass{ujbook}
    \begin{document}
    ...
\end{verbatim}
Save the above as \file{jp2.tex} and run a command.
\begin{verbatim}
    $ uplatex jp2
\end{verbatim}
You will get \file{jp2.dvi}.\bigskip

\upLaTeX\ and \pLaTeX\ (explained later) are often referred to
as (u)\pLaTeX. The output is always a DVI file.
To convert DVI containing Japanese characters into PDF,
the easiest way is to use \emph{dvipdfmx}.
If you have \file{jp2.dvi}, running a command.
\begin{verbatim}
    $ dvipdfmx jp2
\end{verbatim}
will generate \file{jp2.pdf}.\bigskip

Since the default DVI driver in \TL is set to
\Opt{dvips},\footnote{Though dvips supports DVI with Japanese fonts,
additional settings for Ghostscript are required to convert
the resulting PostScript file into PDF.}
it is recommended to add a global driver option
to be passed to all driver-dependent packages
(e.g. \Pkg{graphicx}, \Pkg{color}, \Pkg{hyperref}):
\begin{verbatim}
    \documentclass[uplatex,dvipdfmx]{jsbook}
    \usepackage{graphicx}
    \begin{document}
    ...
\end{verbatim}


% レガシーな pLaTeX は最後
\section{Using legacy \pLaTeX}

Some old Japanese journal classes may require \pLaTeX.
It is a legacy implementation which was born in 1980s,
before the publication of Unicode 1.0.
\pLaTeX\ only supports limited character set JIS~X~0208
(6879 characters).
\begin{verbatim}
    \documentclass{jsbook}% or, \documentclass{jbook}
    \begin{document}
    ...
\end{verbatim}
Save the above as \file{jp3.tex} and run a command.
\begin{verbatim}
    $ platex jp3
\end{verbatim}
You will get \file{jp3.dvi}.


%% 各種文書への導線をはる
\section{Further readings}

There are some documentations in both English and Japanese.

\begin{itemize}
  \item ``The Lua\TeX-ja package'' (\file{luatexja-en.pdf}, in English)
    \begin{itemize}
      \item For beginners, refer to Part I ``User's manual.''
      \item For more interested users, refer to Part II ``Reference.''
      \item For developers, refer to Part III ``Implementations.''
      \item Japanese edition: \file{luatexja-ja.pdf}
    \end{itemize}
  \item ``Guide to \pTeX\ for developers unfamiliar with Japanese''
        (\file{ptex-guide-en.pdf}, in English)
    \begin{itemize}
      \item For developers who aim to support
        \pTeX/\pLaTeX\ and its variants \upTeX/\upLaTeX.
    \end{itemize}
  \item ``About \upLaTeXe''
        (\file{uplatex-en.pdf}, in English)
    \begin{itemize}
      \item For users of \upLaTeX, but a somewhat legacy document.
      \item Japanese edition: \file{uplatex-en.pdf}
    \end{itemize}
  \item ``About \pLaTeXe''
        (\file{platex-en.pdf}, in English)
    \begin{itemize}
      \item For users of \pLaTeX, but a somewhat legacy document.
      \item Japanese edition: \file{platex-en.pdf}
    \end{itemize}
  \item ``Babel-Option \Opt{japanese}''
        (\file{babel-japanese.pdf}, in both English and Japanese)
    \begin{itemize}
      \item Basic introduction of \Opt{japanese} option of \Pkg{babel}.
      \item Provides translated replacement texts for keywords and date.
    \end{itemize}
\end{itemize}


%% 日本の成果物
\section{Japanese deliverables in \TL}

For more details, please refer to individual documentation.

\subsection{Document classes}

Japanese texts can be aligned vertically, called ``tate-gumi'' (縦組).
The horizontal alignment of texts is called ``yoko-gumi'' (横組).
The list below shows \ClsT{tate classes in green}
and \ClsYT{yoko/tate classes in blue} (default is yoko
and class option \Opt{[tate]} enables tate mode).

All the major frameworks for Japanese typesetting provide
``standard class'':
\begin{itemize}
  \item Lua\TeX-ja (\Pkg{luatexja} by Lua\TeX-ja project team)
    \begin{itemize}
      \item Bundle: \Pkg{ltjsclasses}
         (\ClsY{ltjsarticle}, \ClsY{ltjsbook}, \ClsY{ltjsreport})
      \item Bundle: \Pkg{ltjclasses}
         (\ClsY{ltjarticle}, \ClsY{ltjbook}, \ClsY{ltjreport},
          \ClsT{ltjtarticle}, \ClsT{ltjtbook}, \ClsT{ltjtreport})
    \end{itemize}
  \item \upLaTeX\ (\Pkg{uplatex} by \JTDC)
    \begin{itemize}
      \item Bundle: \Pkg{ujclasses}
         (\ClsY{ujarticle}, \ClsY{ujbook}, \ClsY{ujreport},
          \ClsT{utarticle}, \ClsT{utbook}, \ClsT{utreport})
    \end{itemize}
  \item \pLaTeX\ (\Pkg{platex} by \JTDC)
    \begin{itemize}
      \item Bundle: \Pkg{jclasses}
         (\ClsY{jarticle}, \ClsY{jbook}, \ClsY{jreport},
          \ClsT{tarticle}, \ClsT{tbook}, \ClsT{treport})
    \end{itemize}
\end{itemize}

Other major classes:
\begin{itemize}
  \item Class: \ClsYT{jlreq} (by Noriyuki Abe)
    \begin{itemize}
      \item Supports \LuaLaTeX, \upLaTeX\ and \pLaTeX.
      \item This package aims to implement
        ``Requirements for Japanese Text Layout''
        (JLReq, 日本語組版処理の要件, \cite{w3c-jlreq}).
        The class file and necessary JFM (Japanese font metric) files
        are included.
    \end{itemize}
  \item Bundle: \Pkg{jsclasses} (by Haruhiko Okumura \& \JTDC)
    \begin{itemize}
      \item Class: \ClsY{jsarticle}, \ClsY{jsbook}, \ClsY{jsreport}
      \begin{itemize}
        \item Supports only \upLaTeX\ (with option \Opt{[uplatex]}) and \pLaTeX.
        \item In addition to layout adjustments, these classes use
          newly developed JFM set (jis metrics) to avoid strange
          behaviors of standard \pTeX\ JFM set (min10 etc).
          For more details, please refer to \cite{ajt2008okumura}.
        \item Note: when font size other than 10pt is specified,
          the engine primitive \verb+\mag+ is employed.
          Conflicts with other packages can lead to inconsistent
          results in graphics and/or page size.
          To avoid this, specify \Opt{nomag} or \Opt{nomag*} option.
      \end{itemize}
    \end{itemize}
\end{itemize}

\subsection{Packages}

\begin{itemize}
  \item Package: \Pkg{plautopatch} (by Hironobu Yamashita)
    \begin{itemize}
      \item Supports \upLaTeX\ and \pLaTeX, but
        does no harm on other \LaTeX\ (silently ignored).
      \item (u)\pLaTeX\ users are recommended to load this package
        \emph{always}, to resolve errors and incompatibilities with
        Western \LaTeX\ packages which are not aware of (u)\pLaTeX.
    \end{itemize}

\end{itemize}

\subsection{Other support files}

\begin{itemize}
  \item \Pkg{ptex-fontmaps}
    \begin{itemize}
      \item Provides map files to be used with dvipdfmx and dvips.
      \item To customize Japanese fonts to be used in the output,
        please use the command \code{kanji-config-updmap}.
    \end{itemize}
  \item \Pkg{cjk-gs-integrate}
    \begin{itemize}
      \item Supports easy configuration of Ghostscript for
        CJK (Chinese, Japanese and Korean) fonts, which is
        necessary for processing of PostScript file output from dvips.
    \end{itemize}
\end{itemize}

%% 参考文献
\begin{thebibliography}{99}
  \bibitem{w3c-jlreq}
    W3C Working Group,
    \newblock ``{\em Requirements for Japanese Text Layout}''.\\
    \url{https://www.w3.org/TR/jlreq/?lang=en}
  \bibitem{ajt2008okumura}
    Haruhiko Okumura,
    \newblock ``{\em \pTeX\ and Japanese Typesetting}''.
    \newblock The Asian Journal of \TeX, Volume~2, No.~1, 2008.\\
    \url{http://ajt.ktug.org/2008/0201okumura.pdf}
\end{thebibliography}

\end{document}
